Generating Hubbard Model Solutions from Anderson Impurity Model Solutions

### 2021-08-01 notes.tex
added dimer-molecule transformations

>>>>

\section{Solution of the Hubbard dimer using the Anderson molecule}
The Hubbard dimer and Anderson molecules (zero-mode) are defined by the following respective Hamiltonians:
\begin{equation}\begin{aligned}
	H^H &= -t^H\sum_{\sigma}\left(c^\dagger_{1\sigma}c_{2\sigma} + \text{h.c.}\right) + U^H\sum_{i=1,2}\hat n_{i \uparrow}\hat n_{i \downarrow} - \mu^H \sum_{\sigma, i=1,2}\hat n_{i\sigma}\\
	H^A &= -t^A\sum_{\sigma}\left(c^\dagger_{d\sigma}c_{z\sigma} + \text{h.c.}\right) + \epsilon_d^A \sum_{\sigma}\hat n_{d\sigma} + U^A\hat n_{d \uparrow}\hat n_{d \downarrow}
\end{aligned}\end{equation}
In the first Hamiltonian, the indices \(i=1,2\) refer to the two lattice sites that constitute the dimer. In the second Hamiltonian, the subscript \(d\) indicates the impurity site, while the subscript \(z\) indicates the zero-mode site. First, we will assume that the Hubbard dimer is at half-filling (\(\frac{1}{2}U^H = \mu^H\)):
\begin{equation}\begin{aligned}
	\label{hubb_dimer}
	H^H &= -t^H\sum_{\sigma}\left(c^\dagger_{1\sigma}c_{2\sigma} + \text{h.c.}\right) + U^H\sum_{i=1,2}\hat \tau_{i \uparrow}\hat \tau_{i \downarrow} + \left(\frac{1}{2}U^H- \mu^H\right) \sum_{\sigma, i=1,2}\hat n_{i\sigma} + \text{constant}\\
	    &= -t^H\sum_{\sigma}\left(c^\dagger_{1\sigma}c_{2\sigma} + \text{h.c.}\right) + U^H\sum_{i=1,2}\hat \tau_{i \uparrow}\hat \tau_{i \downarrow}
\end{aligned}\end{equation}
Since the Hubbard Hamiltonian is at half-filling, we will also place the impurity at half-filling by setting \(\epsilon_d^A = -\frac{1}{2}U^A\):
\begin{equation}\begin{aligned}
	\label{and_dimer}
	H^A &= -t^A\sum_{\sigma}\left(c^\dagger_{d\sigma}c_{z\sigma} + \text{h.c.}\right) + \left(\epsilon_d^A + \frac{1}{2}U^A\right) \sum_{\sigma}\hat \tau_{d\sigma} + U^A\hat \tau_{d \uparrow}\hat \tau_{d \downarrow} + \text{constant}\\
	    &= -t^A\sum_{\sigma}\left(c^\dagger_{d\sigma}c_{z\sigma} + \text{h.c.}\right) + U^A\hat \tau_{d \uparrow}\hat \tau_{d \downarrow}
\end{aligned}\end{equation}
The first step is to recreate the Hubbard dimer Hamiltonian eq.~\ref{hubb_dimer} using the Anderson molecule Hamiltonian eq.~\ref{and_dimer}:
\begin{equation}\begin{aligned}
	H^H &= -t^H\sum_{\sigma}\left(c^\dagger_{1\sigma}c_{2\sigma} + \text{h.c.}\right) + U^H\sum_{i=1,2}\hat \tau_{i \uparrow}\hat \tau_{i \downarrow}\\
	    &= \frac{1}{2}\left[-t^H\sum_{\sigma}\left(c^\dagger_{1\sigma}c_{2\sigma} + \text{h.c.}\right) + t^H\sum_{\sigma}\left(c^\dagger_{2\sigma}c_{1\sigma} + \text{h.c.}\right)\right] + \frac{1}{2} 2U^H\sum_{i=1,2}\hat \tau_{i \uparrow}\hat \tau_{i \downarrow}\\
	    &= \frac{1}{2}\left[-t^A\sum_{\sigma}\left(c^\dagger_{d\sigma}c_{z\sigma} + \text{h.c.}\right)\bigg\vert_{z \to 2, d \to 1\atop{t^A \to t^H}} + t^A\sum_{\sigma}\left(c^\dagger_{2\sigma}c_{1\sigma} + \text{h.c.}\right)\bigg\vert_{d \to 2,z \to 1\atop{t^A \to t^H}} \right] \\
	    &+ \frac{1}{2} \left(U^A\hat \tau_{i \uparrow}\hat \tau_{i \downarrow}\bigg\vert_{d \to 1\atop{U^A \to 2U^H}} + U^A\hat \tau_{i \uparrow}\hat \tau_{i \downarrow}\bigg\vert_{d \to 2\atop{U^A \to 2U^H}}\right)\\
	    &=\frac{1}{2}\left[H^A\left(t^A \to t^H, U^A \to 2U^H, d \to 1, z \to 2\right) + H^A\left(t^A \to t^H, U^A \to 2U^H, d \to 2, z \to 1\right)\right]
\end{aligned}\end{equation}
The conclusion we can draw from this is that the Hubbard dimer Hamiltonian can be obtained from the Anderson dimer Hamiltonian in the following fashion:
\begin{itemize}
	\item The essential idea is that we have to create a local Hubbard Hamiltonian for each site of the Hubbard lattice by replacing the impurity label \(d\) in the Anderson dimer with the label of the particular site. So if there are two sites, we will get two local Hamiltonians obtained by replacing \(d\) with 1 and 2 respectively. For each local Hamiltonian, the zero-mode label \(z\) is replaced by the site that is nearest to the one that \(d\) is being replaced by. So, if \(d \to 1(2)\), then \(z \to 2(1)\).
	\item This, however, is not the only change that we must make, in order to get the local Hamiltonian for a particular site. Along with \(d\) and \(z\), we must also make the transformations \(t^A \to t^H, U^A \to 2U^H\).
	\item Finally, once we have the local Hamiltonians for sites 1 and 2, we average them to get the total Hubbard Hamiltonian.
\end{itemize}
Note that we expect most of these "rules" to be specific for the dimer, and there will be generalizations to most of them for a general \(N-\)site Hubbard Hamiltonian.

The wavefunctions for the \(N=2\) sector can also be connected through these transformations. Since both the Hamiltonians are analytically solvable, we can write down their groundstate wavefunctions:
\begin{equation}\begin{aligned}
	\ket{\Psi^H_\text{GS}} &= 4t^H\frac{1}{\sqrt 2}\left(\ket{\uparrow_1, \downarrow_2} - \ket{\downarrow_1, \uparrow_2}\right) + \left[\Delta\left( U^H, t^H \right) - U^H\right]{\sqrt 2}\left(\ket{\uparrow_1\downarrow_1, } - \ket{,\uparrow_2\downarrow_2}\right)\\
	\ket{\Psi^A_\text{GS}} &= 4t^A\frac{1}{\sqrt 2}\left(\ket{\uparrow_d, \downarrow_z} - \ket{\downarrow_d, \uparrow_z}\right) + \left[\Delta\left( \frac{1}{2}U^A, t^A \right) - \frac{1}{2}U^A\right]{\sqrt 2}\left(\ket{\uparrow_d\downarrow_d, } - \ket{,\uparrow_z\downarrow_z}\right)\\
	E^H_\text{GS} &=  -\frac{1}{2}\Delta\left( U^H, t^H \right),\quad E^A_\text{GS} =  -\frac{1}{2}\Delta\left( \frac{1}{2}U^A, t^A \right), \quad\Delta(x,y) \equiv \sqrt{x^2 + y^2}\\
\end{aligned}\end{equation}
From the forms of the wavefunctions and eigenenergies, we can immediately write down
\begin{equation}\begin{aligned}
	\ket{\Psi^H_\text{GS}} &= \frac{1}{2}\left[\ket{\Psi^A_\text{GS}}\left(t^A, U^A \to t^H, 2U^H, d \to 1, z \to 2\right) + \ket{\Psi^A_\text{GS}}\left(t^A, U^A \to t^H, 2U^H, d \to 2, z \to 1\right)\right]\\
	E^H_\text{GS} &= E^A_\text{GS}\left(t^A, U^A \to t^H, 2U^H\right)
\end{aligned}\end{equation}
This shows that the rules laid out before work for the Hamiltonians, as well as the wavefunctions and energy eigenvalues of the \(N=2,0,4\) sector. These sectors specifically work because it is only in these sectors can we ensure that \(n_d = n_z\), which is required for the Hubbard Hamiltonian because \(n_1 = n_2\). In the other sectors (\(N=1,3\)), the impurity site and the zero-mode sites have to be singly-occupied in some part, and since the impurity site incurs a single-occupation cost of \(-\frac{U^H}{2}\) which is not borne by the zero-mode site, there is an intrinsic dissimilarity between the two sites of the Anderson molecule in this regime. This dissimilarity does not exist for the Hubbard model, so we cannot hope to connect the two models in this regime.
